---
title: 'Reproducible Research: Peer Assessment 1'
author: "GI"
date: "6/1/2020"
output: html_document
---

## Loading and preprocessing data

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)

# Read in data file
raw.data <- read.csv('activity.csv')
raw.data$date <- as.POSIXct(raw.data$date)

# Create data set that excludes NA values
rm.na.data <- na.omit(raw.data)
```




## What is the mean total number of steps taken per day?

```{r}
# Calculate the total steps taken per day
daily.steps <- aggregate(rm.na.data$steps,by = list(date = rm.na.data$date),FUN = sum)
names(daily.steps)[names(daily.steps) == 'x'] <- 'steps'

# Make histogram of total steps taken each day
ggplot(daily.steps,aes(x = date,y = steps)) + geom_histogram(stat = 'identity',color = 'white')

# Calculate and report the mean and median total steps per day
avg.spd <- mean(daily.steps$steps)
med.spd <- median(daily.steps$steps)
```

The average number of steps taken per day is `r avg.spd`.
The median number of steps taken per day is `r med.spd`.


## What is the average daily activity pattern?

```{r}
# Create subset of 5-min intervals and average number of steps taken
daily.pattern <- aggregate(rm.na.data$steps,by = list(interval = rm.na.data$interval),FUN = mean)
names(daily.pattern)[names(daily.pattern) == 'x'] <- 'avg.steps'

# Make a time series line plot of interval vs avg steps
plot(x = daily.pattern$interval,y = daily.pattern$avg.steps,type = 'l')

# Determine which 5-min interval contains the maximum number of steps
peak.int <- daily.pattern[which.max(daily.pattern$avg.steps),]
peak.int <- peak.int[,1]
```

The `r peak.int` interval contains the maximum number of average steps per day with `r max(daily.pattern$avg.steps)`.


## Imputing missing values

This will be performed by replace all NA values with the average number of steps taken during that 5-minute interval.

```{r}
# Calculate total number of NAs in original dataset
num.na <- sum(is.na(raw.data))

# Replace missing data with average steps for that interval
na.rep.data <- raw.data
na.rep.data$avg.steps <- daily.pattern$avg.steps
na.rep.data$steps[is.na(na.rep.data$steps)] <- na.rep.data$avg.steps

# Calculate total steps taken per day using replaced data set
na.rep.total <- aggregate(na.rep.data$steps,by = list(date = na.rep.data$date),FUN = sum)
names(na.rep.total)[names(na.rep.total) == 'x'] <- 'steps'

# Make histogram of total steps taken each day using replaced data
ggplot(na.rep.total,aes(x = date,y = steps)) + geom_histogram(stat = 'identity',color = 'white')

# Calculate and report the updated mean and median total steps per day
na.rep.avg.spd <- mean(na.rep.total$steps)
na.rep.med.spd <- median(na.rep.total$steps)
```

The average number of steps taken per day using the updated dataset is `r na.rep.avg.spd`.
The median number of steps taken per day using the updated dataset is `r na.rep.med.spd`.

When compared to the original average and median values of `r avg.spd` and `r med.spd`, respectively, we can see that the replacement of NA values had minimal impact on this calculations.


## Are there differences in activity patterns between weekdays and weekends?

```{r}
# Create a new factor in updated data set for Weekday and Weekend
na.rep.data$weekday <- ifelse(weekdays(na.rep.data$date) %in% c("Saturday","Sunday"),'Weekend','Weekday')

# Calculate average steps in 5 minute intervals between Weekday and Weekend
weekday.avg <- aggregate(na.rep.data[na.rep.data$weekday == "Weekday",]$steps,by = list(interval = na.rep.data[na.rep.data$weekday == "Weekday",]$interval),FUN = mean)
weekend.avg <- aggregate(na.rep.data[na.rep.data$weekday == "Weekend",]$steps,by = list(interval = na.rep.data[na.rep.data$weekday == "Weekend",]$interval),FUN = mean)

# Make a panel plot showing the intervals and average steps between Weekday and Weekend
par(mfcol = c(2,1))
plot(weekday.avg$interval,weekday.avg$x,type = 'l',main = 'Weekday Averages',xlab = 'Interval',ylab = 'Avg. Steps')
plot(weekend.avg$interval,weekend.avg$x,type = 'l',main = 'Weekend Averages',xlab = 'Interval',ylab = 'Avg. Steps')
```



